MS797 Bigger Concepts

### 2022-03-05 _academics/ms797-bigger-concepts/ms797-bigger-concepts.Rmd
build

>>>>

## Resampling methods

Resampling methods are used to refine and test models against training data by repeatedly testing against different subsets within the data.

*Cross-validation* can be used to estimate the test error rate of a method (*model assessment*), and *bootstrap* is mostly to measure the accuracy of a parameter or of a method.

### Cross-validation: validation set approach

This approach randomly splits the data into two sets: a *training* set and a *validation* or *test* set. Size ratios seem to be on the order of between 50-50 to 3:1 or 4:1 training-to-test data. The model is fit against the training data, and then `predict()` is used to predict the results for the validation data; the results of `predict` can then be compared against the actual classifications in the validation data to compute the error rate.

This process can be run repeatedly with different splits.

The general process is:

1. Randomly split the dataset into training and validation sets
2. Fit the model to the training data
3. Predict the outcomes of the model on the validation data
4. Convert the predictions into categorical variables that match the field being predicted (like Yes/No, etc)
5. The error rate is the proportion of correct predictions in the validation data.

The code below runs 10 iterations and accumulates the results into a results data frame.

```{r eval=FALSE}
set.seed(1235)
train_pct <- .75

p5_3a.results <- data.frame(iteration = integer(),
                           error_rate = double())

for (i in 1:10) {
  p5_3a.training <- rep(FALSE, nrow(Default))
  p5_3a.training[sample(nrow(Default), nrow(Default) * train_pct)] <- TRUE
  p5_3a.validation <- !p5_3a.training

  p5_3a.fit <-
    glm(
      formula = default ~ income + balance,
      family = binomial,
      data = Default,
      subset = p5_3a.training
    )

  p5_3a.probs <- predict(p5_3a.fit, Default[p5_3a.validation, ], type = "response")
  p5_3a.pred <- ifelse(p5_3a.probs > 0.5, "Yes", "No")
  p5_3a.validation_error_rate <-
    mean(p5_3a.pred == Default[p5_3a.validation, ]$default)
  p5_3a.validation_error_rate_print <-
    round(p5_3a.validation_error_rate, 2)

  p5_3a.results[i,] <- data.frame(
    iteration = i,
    error_rate = 1 - p5_3a.validation_error_rate
  )
}

```

### Cross-validation: leave-one-out (LOOCV)

This method is similar to the validation set approach, except that the validation set is only one element, and the training set is the remaining $n-1$ observations. It is run automatically $n$ times and returns an average of the test errors.

The process is simpler than the validation set approach:

1. Fit the model to the entire data set with `glm()` (omitting the `subset` argument)
2. Run `cv.glm` on the fit
3. The `delta` factor of the return data is the cross-validation error rate for the test data.

Code from the book:

```{r eval=FALSE}
library(boot)
glm.fit <- glm(mpg ~ horsepower, data = Auto)
cv.error <- cv.glm(Auto, glm.fit)
cv.error$delta
```

### Cross-validation: k-fold

K-fold cross-validation is like leave-one-out cross-validation but with a larger test set; it splits the data up into $k$ segments and runs the cross-validation on each. It may give better results if run repeatedly and averaged, as with the validation set approach. It is faster than LOOCV, obviously, because it uses less data.

The process is the same as LOOCV, except a parameter such as `K = 10` is added to the `cv.glm()` call. 5 to 10 is a common range for values of $k$.
